---
title: "Summarizing/Visualizing Simulation Results"
author: "Jennifer Thompson, MPH, and Rameela Chandrasekhar, PhD"
date: "7/14/2018"
output:
  html_document:
    theme: flatly
---

```{r setup, include=FALSE}
library(purrr)
library(dplyr)
library(tidyr)
library(forcats)
library(ggplot2)

## Load and combine simulation results, all stored in results/ in chunks of 250
sim_files <- list.files("results", pattern = "*\\.rds$")
sim_results <- map(paste0("results/", sim_files), readRDS) %>% bind_rows() 

```

## Calculate Quantities for Each Scenario

```{r calc_quants}
## Bias: Mean of betas - true beta
## SE: Mean estimated SEs
## CI width: mean width of the (95%) confidence interval
## Power: % of reps where estimate CI does not include 0
## Coverage: % of reps where estimate CI includes true beta

sim_results2 <- sim_results %>%
  mutate(
    ## "delete" and "impute" are confusing since they both involve imputation;
    ## change to "impute summary" and "impute daily"
    strategy = ifelse(strategy == "delete", "impute summary",
               ifelse(strategy == "impute", "impute daily",
                      strategy)),
    ## Calculate bias
    bias = est_beta - true_beta,
    ## Calculate CI width
    ci_width = ucl - lcl,
    ## Indicators for power, coverage
    power = pvalue < 0.05,
    coverage = true_beta >= lcl & true_beta <= ucl
  )

sim_results_grouped <- sim_results2 %>%
  group_by(miss_type, miss_prop, assoc, strategy, true_beta) %>%
  summarise_at(
    vars(bias, se_beta, ci_width, power, coverage), mean
  ) %>%
  ungroup()

## Create long format for faceted plotting
sim_results_long <- sim_results_grouped %>%
  gather(key = quantity_name, value = quantity_value, bias:coverage) %>%
  ## Create proper factors for plotting
  mutate(
    quantity_f = factor(
      case_when(
        quantity_name == "bias"     ~ 1,
        quantity_name == "se_beta"  ~ 2,
        quantity_name == "ci_width" ~ 3,
        quantity_name == "coverage" ~ 4,
        quantity_name == "power"    ~ 5,
        TRUE ~ as.numeric(NA)
      ),
      levels = 1:5,
      labels = c("Bias", "Std Error", "CI Width", "Coverage", "Power")
    ),
    prop_f = paste0(round(miss_prop * 100), "% Missing"),
    prop_f = fct_reorder(prop_f, miss_prop),
    assoc = factor(
      case_when(
        assoc == "weak"   ~ 1,
        assoc == "mod"    ~ 2,
        assoc == "strong" ~ 3,
        TRUE              ~ as.numeric(NA)
      ),
      levels = 1:3, labels = c("Weak", "Moderate", "Strong")
    ),
    strategy = factor(
      case_when(
        strategy == "ignore" ~ 1,
        strategy == "worst" ~ 2,
        strategy == "impute summary" ~ 3,
        strategy == "impute daily" ~ 4,
        TRUE ~ as.numeric(NA)
      ),
      levels = 1:4,
      labels = c(
        "Ignore (NA = no exp)", "Worst (NA = exp)",
        "Impute summary value", "Impute daily value"
      )
    )
  )

## Dataset of reference lines
## Bias, SE/CI width: 0
## Coverage: 95%
## Power: 80%
sim_refs <- data.frame(
  quantity_f = factor(
    1:5,
    levels = 1:5,
    labels = c("Bias", "Std Error", "CI Width", "Coverage", "Power")
  ),
  ref_value = c(0, 0, 0, 0.95, 0.05)
)

```

```{r function_plot}
plot_miss_type <- function(mtype){
  miss_plot <- ggplot(
    data = subset(sim_results_long, miss_type == mtype)
  ) +
    aes(x = true_beta, y = quantity_value, group = strategy) +
    geom_hline(
      aes(yintercept = ref_value),
      colour = "grey50", linetype = "dotdash", alpha = 0.5, size = 1, data = sim_refs
    ) +
    geom_line(aes(colour = strategy), alpha = 0.5) +
    geom_point(aes(colour = strategy)) +
    labs(title = toupper(mtype)) +
    scale_colour_discrete(name = "Strategy") +
    scale_x_continuous(
      breaks = c(-5, -3, -1, 0),
      minor_breaks = seq(-0.8, -0.2, 0.2),
      name = "True Association, Exposure vs Outcome"
    ) +
    theme_bw() +
    theme(
      axis.title.y = element_blank(),
      legend.position = "bottom",
      legend.direction = "horizontal"
    )
  
  if(mtype == "mcar"){
    miss_plot <- miss_plot + facet_grid(quantity_f ~ prop_f, scales = "free_y")
  } else{
    miss_plot <- miss_plot +
      facet_grid(quantity_f ~ prop_f + assoc, scales = "free_y")
  }
  
  return(miss_plot)
}

save_miss_plot <- function(mtype, mplot, mwidth = 10){
  ggsave(
    filename = sprintf("%s_plot.pdf", mtype),
    mplot,
    path = "results",
    device = "pdf",
    width = mwidth, height = 8, units = "in"
  )
}

```

# MCAR Results

```{r plot_mcar, fig.width = 10}
mcar_plot <- plot_miss_type("mcar")
save_miss_plot(mtype = "mcar", mplot = mcar_plot)
mcar_plot

```

# MAR Results

```{r plot_mar, fig.width = 12}
mar_plot <- plot_miss_type("mar")
save_miss_plot(mtype = "mar", mplot = mar_plot, mwidth = 12)
mar_plot

```

# MNAR Results

```{r plot_mnar, fig.width = 12}
mnar_plot <- plot_miss_type("mnar")
save_miss_plot(mtype = "mnar", mplot = mnar_plot, mwidth = 12)
mnar_plot

```

# General Messages

- "Impute summary value" stinks, don't do it
- Bias increases with more % missing, higher true association regardless (especially for MNAR), but imputing daily values has much less bias than other methods
- Though SEs/CIs are biggest with "impute daily" (not too much worse than ignoring missingness/assuming the best, though)
- Coverage isn't great for any method with >5% missing, especially as true effect sizes increase, but again, "impute daily" still generally the winner.
    - Exception: MNAR, small % missing, with a moderate/strong association with the outcome. There, assuming the worst "wins." This makes sense.
