---
title: Simulating and Assessing Strategies for Summary Variables with Partially Missing
  Data
author: "Jennifer Thompson, MPH"
date: "7/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(purrr)
library(furrr)
library(dplyr)
library(mice)
library(ggplot2)

plan(multiprocess)

```

# Goal

Assess different strategies for summarizing a longitudinal exposure variable
when we are interested in its association with a long-term outcome. In our
motivating clinical example, this is represented by summarizing the amount
(days) of delirium a patient experienced in the hospital and examining its
association with the patient's score on a cognitive assessment performed twelve
months after hospitalization.

## Strategies for Summarizing the Exposure

1. **Ignore It**: Summarize delirium duration by tallying only the delirium
actually observed. Any day with no mental status assessment is considered
non-delirious by omission.

1. **Assume the Worst**: Assume any day with no mental status assessment *is*
delirious, so that the total duration is the sum of all observed delirium + all
missing days.

1. **Deletion**: If any patient-day has no status assessment, we decide that we
cannot determine the exposure value with certainty, and consider the entire
summary value to be missing. Missing *summary* values will be multiply imputed,
using mean severity of illness in the hospital as a covariate; the final model
will be fit using these imputations.

1. **Impute at the Lowest Hierarchy**: We will keep each patient's available
daily data, and impute on a *daily* basis values for days that are missing an
assessment, using that day's severity of illness as a covariate. These imputed
datasets will then be summarized, and the final model will be fit using these
summarized imputed datasets.

We will assess the performance of each strategy under different types and
amounts of missingness, under different associations with the covariate (MAR)
and the exposure (MNAR).

The functions we will use are sourced from the following R scripts, available
in this repository.

```{r source_scripts}
source("introduce_missing.R")
source("summary_strategies.R")
source("fit_models.R")

```

# 1. Simulate Datasets from Motivating Example

Please see `simulate_data.R`.

Result: List of 1000 data.frames, stored as RDS in `analysisdata/`.

```{r read_in_simulated}
simdata_list <- readRDS("analysisdata/simdata.rds")
head(simdata_list[[1]])

## Create list of length(simdata_list) seeds
sim_seeds <- 8675309:(8675309 + (length(simdata_list) - 1))

```

# 2. Simulate Outcomes with Specified Association with Exposure

Please see `simulate_outcome.R`.

Result: List of 1000 lists, each with three elements stored as RDS in
`analysisdata/`. Each element is a data.frame, with the outcome `cogscore`
having a specified association with the exposure `del_actual`.

```{r read_in_outcomes}
outcomedata_list <- readRDS("analysisdata/outcomedata.rds")
head(outcomedata_list[[1]][[1]])

```

We'll start with one simulated dataset and take it through the rest of the
process.

```{r}
df_org <- simdata_list[[1]]
df_out_org <- outcomedata_list[[1]]

```

# 3. Introduce Missingness

We will introduce specific types of missingness to each simulated dataset. For
each scenario, we will set 5%, 20%, 35%, and 50% of individual patient-days to
missing.

- Missing Completely at Random (MCAR)
- Missing at Random (MAR), with the probability of missingness having a weak,
moderate, and strong association with daily severity of illness
- Missing Not at Random (MNAR), with the probability of missingness having a
weak, moderate, and strong association with true daily delirium

We will do this by creating a list of `length(simdata_list)` elements, where
each element is a list of 28 data.frames, all simulated from the seed data.frame
with the specified type/amount of missingness added to the `status_miss`
variable.

```{r introduce_missing, eval = FALSE}
intro_missing <- function(df, seed_set){
  miss_props <- c(0.05, 0.20, 0.35, 0.50)
  strengths <- c("weak", "mod", "strong")
  miss_args <- cross2(miss_props, strengths) %>%
    map(set_names, c("miss_prop", "assoc_strength"))
  
  ## MCAR
  dfs_mcar <- map(miss_props, ~ addmiss_mcar(df = df, seed_set, miss_prop = .))
  
  ## MAR
  dfs_mar <- map2(
    .x = map(miss_args, pluck, "miss_prop"),
    .y = map(miss_args, pluck, "assoc_strength"),
    .f = ~ addmiss_mar(df = df, seed_set, miss_prop = .x, assoc_strength = .y)
  )
  
  ## MNAR
  dfs_mnar <- map2(
    .x = map(miss_args, pluck, "miss_prop"),
    .y = map(miss_args, pluck, "assoc_strength"),
    .f = ~ addmiss_mnar(df = df, seed_set, miss_prop = .x, assoc_strength = .y)
  )
  
  ## Combine into a single list; each data.frame contains columns which specify
  ## information about the missingness
  df_list <- c(dfs_mcar, dfs_mar, dfs_mnar)
  
  return(df_list)
  
}

dfs_missing <- intro_missing(df = df_org, seed_set = 143)

```

# 4. Summarize Exposure, Fit Outcome Model

We will apply each exposure summary strategy to each dataset with missingness,
fit the model appropriately, and extract the estimated beta and standard error.

## Strategy 1: Ignore Missingness

```{r summarize_ignore, eval = FALSE}
tictoc::tic()
mods_ignore <- map(dfs_missing, summarize_ignore) %>%
  map_dfr(fit_lm_typical, dfs_outcome = df_out_org, summary_strat = "ignore")
tictoc::toc()

```

## Strategy 2: Assume the Worst

```{r summarize_worst, eval = FALSE}
tictoc::tic()
mods_worst <- map(dfs_missing, summarize_worst) %>%
  map_dfr(fit_lm_typical, dfs_outcome = df_out_org, summary_strat = "worst")
tictoc::toc()

```

## Strategy 3: Delete It (Then Impute)

```{r summarize_delete, eval = FALSE}
tictoc::tic()
mods_delete <- map(dfs_missing, poss_summarize_delete, seed_set = 1) %>%
  ## Some of our datasets are NULL, because it turns out this method falls apart
  ## when >5% patient-days are missing.
  ## Throw those out, fit the model on the rest.
  discard(is.null) %>%
  map_dfr(fit_lm_impute, dfs_outcome = df_out_org, summary_strat = "delete")
tictoc::toc()

```

## Strategy 4: Impute It On the Daily

```{r summarize_impute, eval = FALSE}
tictoc::tic()
mods_impute <- map(dfs_missing, summarize_impute, seed_set = 1) %>%
  map_dfr(fit_lm_impute, dfs_outcome = df_out_org, summary_strat = "impute")
tictoc::toc()

## One dataset = 1105 seconds. eeek

```

```{r save_mods, eval = FALSE}
## So we don't *have* to run that last one again if we don't want to:
save(mods_ignore, mods_worst, mods_delete, mods_impute,
     file = "analysisdata/fit_mods.Rdata")

```

```{r load_mods}
## If we've already run and saved them:
load("analysisdata/fit_mods.Rdata")

```

Each `mods_xxxx` object is a data.frame, so we just need to combine them:

```{r combine_results}
df_results <- bind_rows(
  mods_ignore, mods_worst, mods_delete, mods_impute
) %>%
  mutate(df_id = 1)

sample_n(df_results, size = 10)

```

Once we do that for all `r length(simdata_list)` simulated datasets, we'll
combine, plot and explore the results!

Note that for a little extra fun, you can use the [`BRRR`](https://github.com/brooke-watson/BRRR) package to let you know when all
that code is done running.